---
title: "epiConv Package"
author: "Li Lin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ISnorm Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}{inputenc}
---

```{r,echo=F,knitr-options,message=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(fig.align = 'center', fig.width = 6, fig.height = 5, dev = 'png')
options(warn=-1)
```

EpiConv is an algorithm to correct batch effects and perform clustering for scATAC-seq data.<br>
To run the scripts below, R package `SingleCellExperiment` is also required. It can be installed through Bioconductor.

We will  integrate resting and LPS-stimulated human bone marrow mononuclear cells (BMMCs, data from [GSE123580](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE123580)).

```{r eval=FALSE}
suppressMessages(library(epiConv))
suppressMessages(library(SingleCellExperiment))

mat<-readMM(file="bmmc_matrix.mtx")
cell_info<-read.table(file="bmmc_ident.tsv")
colnames(cell_info)<-c("barcode","ident")
peak<-read.table(file="peaks.bed")
colnames(peak)<-c("seqnames","start","end")
rownames(mat)<-paste(peak$seqnames,":",
                     peak$start,"-",
                     peak$end,sep="")
colnames(mat)<-cell_info[,1]
sce<-SingleCellExperiment(assays=SimpleList(counts=as(mat,"dgCMatrix")),
                          rowRanges=GRanges(peak),colData=cell_info)
colData(sce)$lib_size<-lib.estimate(assays(sce)$counts)
sce<-subset(sce,select=colData(sce)$lib_size>500)
```
First we read the data and create a SingleCellExperiment object.`lib.estimate` is a simple function to calculate the library size (number of accessible peaks) of single cells. We remove cells with library size <= 500.
```{r eval=FALSE}

row_sample<-1:ncol(sce)
Smat<-run.epiConv(mat=assays(sce)$counts,
                  row_sample=row_sample,
                  lib_size=sce$lib_size,
                  nbootstrap=3,
                  nsample=floor(nrow(sce)*0.5),
                  bin=1000,
                  inf=(-10),
                  backingfile="bmmc.backup",
                  descriptorfile="bmmc.backup.descriptor")
```
We calculate the similarity matrix between cells by `run.epiConv` or the parallel version `run.epiConv.parallel`. The similarity matrix is file-backed. `Smat` is a pointor to the file. A pointor can also be created from existing file `Smat<-attach.big.matrix("bmmc.backup.descriptor")`.<br>
If the data is large, we can down-sample the data by specifying the indices of sampled cells through `row_sample`. The similarities will be calculated between sampled cells and all cells. In the example above, we do not perform down-sampling.


```{r eval=FALSE}
snn_mat<-Smat2snn(Smat,knn=20)
dis<-snn_mat
dis@x<-max(dis@x)-dis@x+1e-9
umap_res<-uwot::umap(X=dis,n_neighbors=20)

```
We calculate the SNN matrix by `Smat2snn` and perform UMAP dimension reduction by `uwot::umap`.

```{r eval=FALSE}
batch<-factor(sce$ident)
Smat_corrected<-deepcopy(Smat,
                         backingfile="bmmc2.backup",
                         descriptorfile="bmmc2.backup.descriptor")
res_anchor<-epiConv.anchor(Smat=Smat_corrected,
                           row_sample=row_sample,
                           batch=batch,
                           reference="Resting",
                           neigs=30,
                           features=NULL,
                           knn_target=50,
                           knn_reference=20,
                           threshold=2)
```
We calculate the knn matrix between datasets by `epiConv.anchor` or the parallel version `epiConv.anchor.parallel`. The knn matrix is stored in `res_anchor$knn_update`.

```{r eval=FALSE}
res_eigs<-epiConv.correct(Smat=Smat_corrected,
                          row_sample=row_sample,
                          batch=batch,
                          reference="Resting",
                          neigs=30,
                          knn_update=res_anchor$knn_update)

```
We perform batch correction using `epiConv.correct` or the parallel version `epiConv.correct.parallel`. `res_eigs` contains the Eigenvectors before and after correction and the Eigenvalues. The similarities between cells in `Smat_corrected` are replaced by corrected values.

```{r eval=FALSE}
snn_mat_corrected<-Smat2snn(Smat_corrected,knn=20)
dis<-snn_mat_corrected
dis@x<-max(dis@x)-dis@x+1e-9
umap_res_corrected<-uwot::umap(X=dis,n_neighbors=20)

plot(umap_res,pch="+",cex=0.5,col=factor(sce$ident))
plot(umap_res_corrected,pch="+",cex=0.5,col=factor(sce$ident))
```
Finally, we perfrom UMAP dimension reduction on corrected data and compare the results before and after correction.

```{r eval=FALSE}
clust<-epiConv.louvain(snn=snn_mat_corrected,resolution=c(0.8,0.6,0.4,0.2))
head(clust)
```
Louvain clustering can also be performed on the SNN matrix.
